Scenario 2 - MPX in the EU

### 2022-11-07 MS3.Rmd
11/6/22 Upload

@@

### Milestone 3: 1. Sub-setting Rows and Columns as Needed- Clean individual Data Frames

Data Frame Cleaning

@@

#Cleaning the denominators df
euro_pop_denominators <- euro_pop_denominators %>% 
  filter( TIME_PERIOD == 2022) %>%
  rename( CountryCode = geo , total_pop_2022 = 'OBS_VALUE') %>%
  select( CountryCode, total_pop_2022)

@@

### Milestone 3: 2. Summarizing Data and Creating New Vectors, Organize Vectors for Use

@@

#We now add the EU_Region vector based on the country name.
euro_mpx_cases <- euro_mpx_cases %>% 
  mutate(EU_region = case_when(
    CountryCode %in% sr_eastern_e ~ "Eastern Europe",
    CountryCode %in% sr_northern_e ~ "Northern Europe",
    CountryCode %in% sr_southern_e ~ "Southern Europe",
    CountryCode %in% sr_western_e ~ "Western Europe",
    CountryCode %in% sr_western_a ~ "Western Asia"
    ))
```

### Milestone 3: 3. Clean Variables and Combined Data-sets

#Combined the denominations data for 2022 pop with euro_mpx_cases dataset

euro_mpx_cases <-  inner_join( euro_pop_denominators, euro_mpx_cases, 
                                by= "CountryCode")
#Join census data
euro_mpx_cases <- inner_join(euro_census_stats, euro_mpx_cases,
                              by= "CountryCode")
#Rearrange column order
euro_mpx_cases <- relocate(euro_mpx_cases, floor_month, EU_region, 
                          total_pop_2022 , cases_grouped_monthly, 
                          .before = CountryCode)

euro_mpx_cases<- relocate(euro_mpx_cases, CountryCode, .before = total_pop_2022)

@@

### Milestone 3: 4. Descriptive Stats of EU Data and Data Dictionary

Descriptive stats:
We find that the census information for 7 of the 29 countries in the EU are not
available. These "0" values were replaced with "NA" to prevent graphing issues.

Region Data:
We see that the regions in EU include various numbers of countries, thus having unequal
populations.

"Eastern Europe" : 6 countries, total region population = 89171711   
"Northern Europe": 9 countries, total region population = 37518701   
"Southern Europe": 7 countries, total region population = 133879004   
"Western Asia"   : 1 countries, total region population = 904705   
"Western Europe" : 6 countries, total region population = 190251495

@@

visual <- ggplot(data = euro_mpx_cases, 
                 aes(x = EU_region , y = cases_grouped_monthly))+ 
  theme_minimal(base_size = 15)

visual

```{r}
#Upload the Data Dictionary for the central data-frame: "euro_mpx_cases"
Data_Dictionary <- read_excel("~/PHW251_Fall2022/251/MS3_data_dictionary.xlsx")
Data_Dictionary

>>>>

Group R: Team Logistics Agreement

9/19/22

1.	Project option selection

Scenario 2: Monkeypox (MPX) infection rates in the European Union

2.	Create git repository

https://github.com/dcoblin/251.git

3.	Roles and responsibilities

a.	Team's preferred communication method

WhatsApp

b.	Team's preferred meeting times and frequency

Weekends/Post 6 PM on weekdays

c.	Team's preferred method for tracking progress

Google docs

d.	Point person for contacting course facilitators with questions

Jose/Dawson

e.	Point person for submitting all milestones

Dawson/Jose

@@

## Milestone 3: 1. Sub-setting Rows and Columns as Needed- Clean individual Data Frames

### Data Frame Cleaning

@@

#Cleaning the denominators df
euro_pop_denominators <- euro_pop_denominators %>% 
  filter( TIME_PERIOD == 2022) %>%
  rename( CountryCode = geo , country_pop_2022 = 'OBS_VALUE') %>%
  select( CountryCode, country_pop_2022)

@@

## Milestone 3: 2. Summarizing Data and Creating New Vectors, Organize Vectors for Use

@@

#We now add the EU_Region vector based on the country name.
euro_mpx_cases <- euro_mpx_cases %>% 
  mutate(EU_region = case_when(
    CountryCode %in% sr_eastern_e ~ "Eastern Europe",
    CountryCode %in% sr_northern_e ~ "Northern Europe",
    CountryCode %in% sr_southern_e ~ "Southern Europe",
    CountryCode %in% sr_western_e ~ "Western Europe",
    CountryCode %in% sr_western_a ~ "Western Asia"
    ))

@@

## Milestone 3: 3. Clean Variables and Combined Data-sets

@@

euro_mpx_cases <-  inner_join( euro_pop_denominators, euro_mpx_cases, 
                                by= "CountryCode")
#Join census data.
euro_mpx_cases <- inner_join(euro_census_stats, euro_mpx_cases,
                              by= "CountryCode")
#Rearrange column order.
euro_mpx_cases <- relocate(euro_mpx_cases, EU_region, floor_month, 
                          country_pop_2022 , cases_grouped_monthly, 
                          .before = CountryCode)

#Combined the denominations data for 2022 pop with euro_mpx_cases dataset.
sum_region_cases <- euro_mpx_cases %>% group_by(EU_region) %>%
  summarise(sum(cases_grouped_monthly))

#Join the numeric sums into the main dataframe.
euro_mpx_cases <- inner_join(sum_region_cases, euro_mpx_cases,
                              by= "EU_region")

#Relocate the columns to a adjust fluidity of the data.
euro_mpx_cases<- relocate(euro_mpx_cases, floor_month, CountryCode, 
                          country_pop_2022, cases_grouped_monthly, 
                          .before = EU_region)
#Rename the sum of region cases that are confirmed.
euro_mpx_cases<- euro_mpx_cases %>%
   rename( sum_region_cases = `sum(cases_grouped_monthly)`)
```

## Milestone 3: 4. Descriptive Stats of EU Data and Data Dictionary

### Descriptive stats:
We find that the census information for 7 of the 29 countries in the EU are not
available. These "0" values were replaced with "NA" to prevent graphing issues.

### Region Data:
We see that the regions in EU include various numbers of countries, thus having 
unequal populations. A useful metric to create a graph will be the calculation 
of cases per 100,000. This will allow us to use 1 categorical variable (EU Region)
to a quantitative variable (number of confirmed cases)

We compiled the relevant confirmed case information per region in the EU.

### The pertinent information per EU region is listed below:
```{r echo=FALSE}
Table_1 <- tibble(
  EU_Region = c("Eastern Europe", "Northern Europe", "Southern Europe",
               "Western Asia", "Western Europe" ),  
  Countries_per_Region = c(6, 9, 7, 1, 6),
  Total_Region_Pop = c(89171711, 37518701,133879004, 904705, 190251495),
  Total_cases_per_region = c( 217, 619, 7958, 4, 8214)
  )

Table_1
```
## Data Dictionary

```{r}
#Upload the Data Dictionary for the central data-frame: "euro_mpx_cases"
Data_Dictionary <- read_excel("~/PHW251_Fall2022/251/MS3_data_dictionary.xlsx")
Data_Dictionary
```
